Concurrent Programming with Harmony

### 2020-07-30 book/paper.tex
Fixed typo

There is no critical section.  If one process is executing instructions
very slowly, this does not negatively impact the other processes, as it
would with solutions based on critical sections.
On the contrary, it helps them because it creates less contention.
Unfortunately, the solution is not practical for the following reasons:
\begin{itemize}
\item The \textit{items} array must be of infinite size if an unbounded number
of items may be produced;
\item Each slot in the array is only used once, which is inefficient;
\item the \texttt{inc} and \texttt{exch} interlock instructions are not
universally available on existing processors.
\end{itemize}
However, in the literature there are many examples of practical
non-blocking (aka \emph{wait-free})
\index{wait-free synchronization}%
synchronization algorithms~\cite{}.
% TODO

@@

\section*{Exercises}
\begin{problems}
\item See if you can implement barrier synchronization for \texttt{N} processes
with just three semaphores.  Busy waiting is not allowed.  How about just two?
(As always, the Little Book of Semaphores~\cite{Downey09} is a good resource
for solving semaphore problems.  Look for the double turnstile solution.)
\item Implement barrier synchronization with Mesa condition variables.
(You may want to use a double turnstile approach here as well.)
\end{problems}

@@

In each case, $s$ must be a set.
For list comprehensions, the list will be sorted by $e$.
For dictionary comprehensions, each $e$ will be mapped to \texttt{f}($e$).

@@

\vspace{1em}
\begin{tabular}{|l|l|}
\hline
\textit{lv} = [lv =]... $e$ & \textit{lv} is an lvalue and $e$ is an expression\\
\hline
\textit{lv} \texttt{[op]=} $e$ & \texttt{op} is one of \texttt{+}, \texttt{-},
\texttt{*}, \texttt{/}, \texttt{//}, \texttt{\%},
\texttt{\string\&}, \texttt{|}, \texttt{\string\^},
\texttt{and}, and \texttt{or}\\
\hline
\texttt{pass} & do nothing\\
\hline
\texttt{del} \textit{lv} & delete\\
\hline
\texttt{assert $b$ [, $e$]} & $b$ is a boolean.  Optionally report value of expression $e$\\
\hline
\texttt{const a = $e$} & \texttt{a} is a bounded variable, $e$ is a constant expression\\
\hline
\texttt{def m $a$: S} & \texttt{m} is an identifier, $a$ a bounded variable, \texttt{S} a list of statements\\
\hline
\texttt{let $a$ = $e$ [let ...]: S} & $a$ is a bounded variable, $e$ is an expression, \texttt{S} a list of statements\\
\hline
\texttt{if $b$: S} & $b$ is a boolean, \texttt{S} a list of statements\\
\hline
\texttt{while $b$: S} & $b$ is a boolean, \texttt{S} a list of statements\\
\hline
\texttt{for $a$ in $e$: S} & $a$ is a bounded variable, $e$ is a set,
                            \texttt{S} a list of statements\\
\hline
\texttt{atomic: S} & \texttt{S} a list of statements\\
\hline
\texttt{spawn m $e$ [, $t$]} & \texttt{m} is a method,
$e$ is an expression, $t$ is a tag (an expression) \\
\hline
\texttt{trap m $e$} & \texttt{m} is a method and $e$ is an expression \\
\hline
\texttt{go $c$ $e$} & $c$ is a context, $e$ is an expression \\
\hline
\texttt{import m} & \texttt{m} identifies a module \\
\hline
\end{tabular}
\vspace{1em}

Harmony does not support Python-style iterators.
To iterate through a dictionary, list, or tuple, use:
\texttt{for $v$ in keys($e$):~S}.

@@

\vspace{1em}
\begin{tabular}{|l|l|}
\hline
\texttt{subseq($t$, $b$, $b$)} & return a \emph{slice} of list $t$ starting
at index $b$ and ending just before $f$\\
\hline
\texttt{append($t$, $e$)} & append $e$ to list $t$\\
\hline
\texttt{head($t$)} & return the first element of list $t$\\
\hline
\texttt{tail($t$)} & return all but the first element of list $t$\\
\hline
\texttt{reversed($t$)} & reverse a list \\
\hline
\texttt{sorted($t$)} & sorted set or list \\
\hline
\texttt{set($t$)} & convert values of a dict or list into a set \\
\hline
\texttt{list($t$)} & convert set into a list \\
\hline
\texttt{values($t$)} & convert values of a dict into a list sorted by key \\
\hline
\texttt{items($t$)} & convert dict into (key, value) list sorted by key \\
\hline
\texttt{enumerate($t$)} & like Python enumerate \\
\hline
\texttt{sum($t$)} & return the sum of all elements in $t$\\
\hline
\texttt{list2bag($t$)} & convert list $t$ into a bag \\
\hline
\texttt{listQsort($t$)} & sort list $t$ using quicksort\\
\hline
\end{tabular}

@@

\chapter{Harmony Language Details}

>>>>

There is no critical section.  If one process is executing instructions
very slowly, this does not negatively impact the other processes, as it
would with solutions based on critical sections.
On the contrary, it helps them because it creates less contention.
Unfortunately, the solution is not practical for the following reasons:
\begin{itemize}
\item The \textit{items} array must be of infinite size if an unbounded number
of items may be produced;
\item Each slot in the array is only used once, which is inefficient;
\item the \texttt{inc} and \texttt{exch} interlock instructions are not
universally available on existing processors.
\end{itemize}
However, in the literature there are many examples of practical
non-blocking (aka \emph{wait-free})
\index{wait-free synchronization}%
synchronization algorithms.
% TODO~\cite{}.

@@

\section*{Exercises}
\begin{problems}
\item See if you can implement barrier synchronization for \texttt{N} processes
with just three semaphores.  Busy waiting is not allowed.  How about just two?
(As always, the Little Book of Semaphores~\cite{Downey09} is a good resource
for solving synchronization problems with semaphores.
Look for the \emph{double turnstile}
\index{double turnstile}%
solution.)
\item Implement barrier synchronization with Mesa condition variables.
(You may want to use a double turnstile approach here as well.)
\end{problems}

@@

In each case, $s$ must be a set or a dictionary.
\autoref{ap:details} provides more information on comprehensions.

@@

\vspace{1em}
\begin{tabular}{|l|l|}
\hline
\textit{lv} = [lv =]... $e$ & \textit{lv} is an lvalue and $e$ is an expression\\
\hline
\textit{lv} \texttt{[op]=} $e$ & \texttt{op} is one of \texttt{+}, \texttt{-},
\texttt{*}, \texttt{/}, \texttt{//}, \texttt{\%},
\texttt{\string\&}, \texttt{|}, \texttt{\string\^},
\texttt{and}, and \texttt{or}\\
\hline
\texttt{pass} & do nothing\\
\hline
\texttt{del} \textit{lv} & delete\\
\hline
\texttt{assert $b$ [, $e$]} & $b$ is a boolean.  Optionally report value of expression $e$\\
\hline
\texttt{const a = $e$} & \texttt{a} is a bounded variable, $e$ is a constant expression\\
\hline
\texttt{def m $a$: S} & \texttt{m} is an identifier, $a$ a bounded variable, \texttt{S} a list of statements\\
\hline
\texttt{let $a$ = $e$ [let ...]: S} & $a$ is a bounded variable, $e$ is an expression, \texttt{S} a list of statements\\
\hline
\texttt{if $b$: S} & $b$ is a boolean, \texttt{S} a list of statements\\
\hline
\texttt{while $b$: S} & $b$ is a boolean, \texttt{S} a list of statements\\
\hline
\texttt{for $a$ in $e$: S} & $a$ is a bounded variable, $e$ is a set,
                            \texttt{S} a list of statements\\
\hline
\texttt{atomic: S} & \texttt{S} a list of statements\\
\hline
\texttt{spawn m $e$ [, $t$]} & \texttt{m} is a method,
$e$ is an expression, $t$ is a tag (an expression) \\
\hline
\texttt{trap m $e$} & \texttt{m} is a method and $e$ is an expression \\
\hline
\texttt{go $c$ $e$} & $c$ is a context, $e$ is an expression \\
\hline
\texttt{import m} & \texttt{m} identifies a module \\
\hline
\end{tabular}

@@

\vspace{1em}
\begin{tabular}{|l|l|}
\hline
\texttt{subseq($t$, $b$, $f$)} & return a \emph{slice} of list $t$ starting
at index $b$ and ending just before $f$\\
\hline
\texttt{append($t$, $e$)} & append $e$ to list $t$\\
\hline
\texttt{head($t$)} & return the first element of list $t$\\
\hline
\texttt{tail($t$)} & return all but the first element of list $t$\\
\hline
\texttt{reversed($t$)} & reverse a list \\
\hline
\texttt{sorted($t$)} & sorted set or list \\
\hline
\texttt{set($t$)} & convert values of a dict or list into a set \\
\hline
\texttt{list($t$)} & convert set into a list \\
\hline
\texttt{values($t$)} & convert values of a dict into a list sorted by key \\
\hline
\texttt{items($t$)} & convert dict into (key, value) list sorted by key \\
\hline
\texttt{enumerate($t$)} & like Python enumerate \\
\hline
\texttt{sum($t$)} & return the sum of all elements in $t$\\
\hline
\texttt{list2bag($t$)} & convert list $t$ into a bag \\
\hline
\texttt{listQsort($t$)} & sort list $t$ using quicksort\\
\hline
\end{tabular}

@@

\chapter{Harmony Language Details}
\label{ap:details}
